Conditional selectivity for Hexastores

### 2011-07-20 conditional/trunk/selectivity-hexastore.tex
reindent
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@15 43ef1190-90fd-4dd7-a769-d0121f02b80e
make sense of matching
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@16 43ef1190-90fd-4dd7-a769-d0121f02b80e

\newcommand{\rdfnode}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

@@

It is easy to point out cases where the independence assumption is
unlikely to hold: If a predicate is \rdfnode{rdf:type}, then only
classes are likely objects, and their frequency  would likely be
different from e.g. literals.

@@

Previous authors have taken selectivity as analogous to the
probability of matching a SPARQL node\todo{terminology, (SPARQL nodes
  vs. RDF terms) could be better} or triple with a node or triple
in the data. For a triple, $<S, P, O>$ the selectivity is the
intersection of the selectivity of each node.

From elementary statistics, we know that the intersection probability
can be written as
\begin{equation}
Pr(A_1 \cap A_2  \cap \ldots \cap A_n) = 
  Pr(A_n | A_1 \cap A_2  \cap \ldots \cap A_{n-1}) \\
  \cdot Pr(A_{n-1} | A_1 \cap A_2  \cap \ldots \cap A_{n-2}) 
  \cdot \ldots \cdot Pr(A_2 | A_1) \cdot Pr(A_1) ,
\end{equation}
where the $|$ symbol is read as ``given'', e.g. ``given a predicate
\rdfnode{rdf:type} what is the probability that the object is \rdfnode{foaf:Person}?''

@@

Now, given a query, we shall explain how the selectivity can be
computed exactly rather than being estimated. First, we define a
function $cm(x)$ which returns the number of matches of a given SPARQL
node $x$. \todo{needs to be a lot more precise}

>>>>

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

@@

It is easy to point out cases where the independence assumption is
unlikely to hold: If a predicate is \rdfterm{rdf:type}, then only
classes are likely objects, and their frequency  would likely be
different from e.g. literals.

@@

Previous authors have taken selectivity as analogous to the
probability of matching a SPARQL term (RDF Term or variable) with an
RDF term in the data or a SPARQL triple pattern with a triple in the
data.  For a triple, $<S, P, O>$ the selectivity is the intersection
of the selectivity of each node.

From elementary statistics, we know that the intersection probability
can be written as
\begin{equation}
Pr(A_1 \cap A_2  \cap \ldots \cap A_n) = 
  Pr(A_n | A_1 \cap A_2  \cap \ldots \cap A_{n-1}) \\
  \cdot Pr(A_{n-1} | A_1 \cap A_2  \cap \ldots \cap A_{n-2}) 
  \cdot \ldots \cdot Pr(A_2 | A_1) \cdot Pr(A_1) ,
\end{equation}
where the $|$ symbol is read as ``given'', e.g. ``given a predicate
\rdfterm{rdf:type} what is the probability that the object is \rdfterm{foaf:Person}?''

@@

Now, given a query, we shall explain how the selectivity can be
computed exactly rather than being estimated. First, we define a
function $cm(x)$ which returns the number of matches of a given SPARQL
term $x$. \todo{needs to be a lot more precise}
